---
title: "Box Plots"
format: html
engine: knitr
filters:
  - webr
---

```{r}
#| label: loading data sets
#| include: false 
library(tidyverse)

#penguin data
library(palmerpenguins)
penguins<- palmerpenguins::penguins

#Chick weight data
data("ChickWeight")
chickWeight<-ChickWeight

#president data
data("presidents")
presidents<-presidents
```

## What is a box plot

A box plot is a graph that shows a ***numeric*** summary (y-axis) of either a ***categorical or numerical*** variable (x-axis). It gives information on the spread of data.

Here is the framework of a box plot, with a normally distributed data set.

```{r}
#| label: generating frame plot
#| warning: false
#| echo: false
data <- rnorm(100, mean = 50, sd = 10)

ggplot(data.frame(Value = data), aes(y = Value)) +
  geom_boxplot() +
  labs(title = "Sample Box and Whisker Plot", y = "Numeric continous", x = " Categorical ordinal/nominal")
```

The "**box**," as shown below, represents three values: the 25th percentile, median (50th percentile), and the 75th percentile. Each horizontal bar corresponds to a number where that percentage of the data falls below. In other words, the horizontal bar closest to the x-axis represents where 25%\ of the data falls below, the bar in between the box is the median, or where 50%\ of the data fall below, and the furthest horizontal bar represents where 75%\ of the data falls below.

The vertical lines that emerge from the box are called **whiskers**. The whisker closest to the x-axis represents the minimum calculated by subtracting three times the interquartile range (75th percentile-25th percentile) from the value for the 25th percentile. The furthest whisker from the x-axis represents the maximum calculated by subtracting three times the interquartile range from the 75th percentile.

Data points that exist outside of these calculated minimum and maximum values exist as points separated from the plot, called outliers.

## Penguin Data

Using the penguin dataset, we can compare the flipper lengths (variable: flipper_length_mm) of the different species (variable: species) of penguins.

*Note:* We will call on the geom_boxplot( ) function when we code using ggplot

```{r}
#| warning: FALSE 

ggplot(data=penguins, mapping = aes(x=species, y=flipper_length_mm))+
  geom_boxplot()+
  labs(x="Species of Antarctic penguin", y="Flipper length (mm)", title = "Species of Antarctic penguins have different flipper lengths")
```

For the Adelie penguins, there are outliers on either side of the whiskers. These are values fall short/exceed the whisker value calculation.

Now, compare the flipper lengths (variable: flipper_length_mm) of the different sexes (variable: sex) of penguins


