---
title: "Histograms"
format: html
engine: knitr
filters:
  - webr
---

```{r}
#| label: loading data sets
#| include: false 
library(tidyverse)

#penguin data
library(palmerpenguins)
penguins<- palmerpenguins::penguins

#Chick weight data
data("ChickWeight")
chickWeight<-ChickWeight

#president data
data("presidents")
presidents<-presidents
```

## What is a histogram?

A histogram is a graph that compares the ***frequency*** (y-axis) of any ***categorical or numeric*** variable (x-axis). In other words, it measures how often a value from the x-axis appears in a data set.

Here is the framework of a histogram:

```{r}
#| label: generating frame plot
#| warning: false
#| echo: false
plot(x = 1,                 
     xlab = "Categorical/numeric variable", 
     ylab = "Frequency",
     xlim = c(0, 100), 
     ylim = c(0, 100),
     main = "Histogram framework",
     type = "n")
```

## Penguin Data

In the penguin dataset, we can use a histogram to determine how often each bill depth (variable: bill_depth_mm) value appears in the dataset.

Here, there is a deviation from the ggplot framework in the home page. "bins" is used to tell R how many bars to create in your histogram. You can change this number as needed. Ideally, you want to keep the bin number large enough to see patterns in your data , but small enough to keep outliers to a minimum.

```{r, warning=FALSE}
ggplot(data=penguins, mapping=aes(x=bill_depth_mm, na.rm=TRUE))+
  geom_histogram(bins=20)+
  labs(title="Distribution of Antarctic Penguin Bill Depth", x="Bill depth (mm)", y="Frequency")

```

Now, create a histogram to see the distribution of chick weights (variable "weight") using the chickWeight dataset.


