---
title: "Histograms"
format: html
engine: knitr
filters:
  - webr
include-in-header:
  file: checker.js
---

```{r}
#| include: false
source("checker.R")
```

```{r}
#| label: loading data sets
#| include: false 

library(tidyverse)

#penguin data
library(palmerpenguins)
penguins<- palmerpenguins::penguins

#Chick weight data
data("ChickWeight")
ChickWeight<-ChickWeight

```

## What is a histogram?

A histogram is a graph that compares the ***frequency*** (y-axis) of any ***categorical or numeric*** variable (x-axis). In other words, it measures how often a value from the x-axis appears in a data set.

Here is the framework of a histogram:

```{r}
#| label: generating frame plot
#| warning: false
#| echo: false
plot(x = 1,                 
     xlab = "Categorical/numeric variable", 
     ylab = "Frequency",
     xlim = c(0, 100), 
     ylim = c(0, 100),
     main = "Histogram framework",
     type = "n")
```

## Penguin Data

In the penguin dataset, we can use a histogram to determine how often each bill depth (variable: bill_depth_mm) value appears in the dataset.

Here, there is a deviation from the ggplot framework in the home page. "bins" is used to tell R how many bars to create in your histogram. You can change this number as needed. Ideally, you want to keep the bin number large enough to see patterns in your data, but small enough to keep outliers to a minimum.

```{r, warning=FALSE}
ggplot(data=penguins, mapping=aes(x=bill_depth_mm, na.rm=TRUE))+
  geom_histogram(bins=20)+
  labs(title="Distribution of Antarctic Penguin Bill Depth", x="Bill depth (mm)", y="Frequency")

```

This data is pretty noisy. There appears to be three peaks. There are three species of penguins in this data set. Perhaps each peak represents each species?

Let's see if there is a trend in **Gentoo** penguins.

First, we must filter the penguin data to include only Gentoo penguins by creating a new variable, "gentooPenguins".

```{r}
#| include: false
gentooPenguins<-penguins %>%
  filter(species == "Gentoo")
```

```{webr-r}
#| exercise: ex1
___<-penguins %>%
  filter(species == "___")
```

Now let's create a histogram of the distribution of Gentoo penguin bill depths. Don't worry about adding a title or axes labels!

::: panel-tabset
## Code editor

```{webr-r}
ggplot(data=___, mapping=aes(x=___, na.rm=TRUE))+
  geom_histogram(bins=20)

```

## Hint

R is case sensitive! The data set is named "gentooPenguins" and the variable is "bill_depth_mm".

## Answer

```{r}
#| eval: false
ggplot(data=gentooPenguins, mapping=aes(x=bill_depth_mm, na.rm=TRUE))+
  geom_histogram(bins=20)
```

This data still appears to show no clear pattern. Perhaps there is no trend in bill depth for Gentoo penguins in the data provided.
:::

## Chick Data

Let's try using the ChickWeight data set! Using the model code above, create a histogram to see the distribution of chick weights (variable "weight") using the ChickWeight dataset.

::: panel-tabset
## Code editor

```{webr-r}
#Type your code here
```

## Hint

R is case sensitive! The data set is named "ChickWeight" and the variable is "weight".

## Answer

```{r}
#| eval: false
ggplot(data=ChickWeight, mapping=aes(x=weight, na.rm=TRUE))+
  geom_histogram(bins=20)
```
:::

How is the data distributed? Are there more chicks that fall into a weight interval to the **left** of the graph, to the **right** of the graph, or is it **normally distributed**?

The histogram is right skewed, meaning throughout the data, most chicks' weights fall between 0-100g.
