---
title: "Tips and Tricks"
format: html
engine: knitr
filters:
  - webr
---

```{r}
#| label: loading tidyverse
#| include: false 
library(tidyverse)
library(palmerpenguins)
```

Here are some tools that can improve your data visualizations

**Colors** can be used as a tool to emphasize an idea on a graph. To change the color of the points/lines on your graph, select the color you want using \*color="\_"\* in the function specifying the graph type. For example, to change the default black points in a scatter plot to blue,

``` r
ggplot(data=dataset, mapping=aes(x=xVariable, y=yVariable))+ 
geom_point(color="blue"))
```

There are many colors to choose from, but it is important to consider how your audience would perceive them. Are they bright and straining? Are they accessible to someone who has visual impairments?

**Facet wrap** is a function that can display multiple data visualizations at once. It is used to show graphs with the same axes with a change in a categorical variable. For example, we can compare the weights of all three Antarctic penguin species in one data visualization. It's useful when trying to compare data, without creating a visually busy graph.

To do this, call on the facet_wrap( ) function and inside the parentheses, specify the categorical variable using a tilde (~).

```{r}
ggplot(data=penguins, mapping=aes(x=bill_length_mm, y=body_mass_g))+ 
  geom_point()+ 
  facet_wrap(~species)+ 
  labs(title="Bill length by body mass for Antarctic Penguin Species")

```