---
title: "Bar Plots"
format: html
engine: knitr
filters:
  - webr
---

```{r}
#| label: loading data sets
#| include: false 
library(tidyverse)

#penguin data
library(palmerpenguins)
penguins<- palmerpenguins::penguins

#Chick weight data
data("ChickWeight")
chickWeight<-ChickWeight

#president data
data("presidents")
presidents<-presidents
```

## What is a bar plot?

A bar plot compares a categorical variable (x-axis) to a numeric continuous variable (y-axis). Each bar represents the average *or* the frequency for each categorical variable. The bars can also be used to represent the counts, or number of occurences within a categorical variable.

Here is the framework for a bar graph:

```{r}
#| label: generating frame plot
#| warning: false
#| echo: false

```

## Penguin data
Now, let's see how the average bill depths of female and male Adelie penguins compare, using the penguins dataset.

*Note:* We will call on the geom_bar( ) function using ggplot. When we call this function, we use stat="identity" to tell R we want our y-axis to represent the average bill depth, instead of counting how many sexes are represented for Adelie penguins in this data set.

```{r}
#first, we must group_by( ) and summarize ( ) the data on Adelie penguins
adeliePenguins<- penguins |>
  filter(species=="Adelie")|>
  group_by(sex, na.rm=TRUE) |>
  summarize(avgBillDepth = mean(bill_depth_mm, na.rm=TRUE), sdBillDepth = sd(bill_depth_mm, na.rm=TRUE))
  
ggplot(data=adeliePenguins, mapping = aes(x = sex, y=avgBillDepth, fill = sex))+
  geom_bar(stat="identity")
```

We can also add error bars to our graphs when we use them to represent averages. Error bars help represent the spread of data using a quantity known as the standard deviation. The standard deviation represents, on average, how much every data point is from the mean.

The error bars extend from the top of the column plus and minus the standard deviation.

Here is the same plot, with the error bars.

*Note*: To create error bars, call on the geom_errorbar( ) function, with a nested mapping=aes( ), as shown

```{r}
ggplot(data=adeliePenguins, mapping = aes(x = sex, y=avgBillDepth, fill = sex))+
  geom_bar(stat="identity")+
  geom_errorbar(mapping=aes(ymin=avgBillDepth-sdBillDepth, ymax=avgBillDepth+sdBillDepth))
```
