---
title: "Scatter Plots"
format: html
engine: knitr
filters:
  - webr
---

```{r}
#| label: loading data sets
#| include: false 
library(tidyverse)

#penguin data
library(palmerpenguins)
penguins<- palmerpenguins::penguins

#Chick weight data
data("ChickWeight")
chickWeight<-ChickWeight

#president data
data("presidents")
presidents<-presidents
```

## What is a scatter plot?

A scatter plot compares both a numerical values on the x and y axes. They can be used to show trends over time, or to show the correlation between two variables.

Here is the plot framework:

```{r}
#| label: generating frame plot
#| warning: false
#| echo: false
plot(x = 1,                 
     xlab = "Numeric variable", 
     ylab = "Numeric variable",
     xlim = c(0, 100), 
     ylim = c(0, 100),
     main = "Scatter plot framework",
     type = "n")
```

## Penguin Data

In the penguin data set, we can examine the relationship between bill depth and bill length.

*Note*: To generate a scatter plot, we call on the geom_point( ) function with ggplot.

```{r, warning = FALSE}
ggplot(data=penguins, mapping=aes(x=bill_length_mm, y=bill_depth_mm))+
  geom_point()
```

## Line of best fit

So we have these dots, but how do we know if one variable *really* is related to another? We can use the Pearson correlation coeffcient, a measure of the linear relationship of two numeric variables. A perfect direct correlation would have a slope of 1, and a perfect indirect correlation would have a slope of -1.

In order to perform regression analysis, or how the x and y variables are related, we can use a linear model equation y=mx + b, where y is a function of the change in a given y-value times its respective x-value, plus the y-intercept, or where the y-value crosses the y=0 axis.

```{r, warning=FALSE}
billDepthAndLengthModel=lm(data=penguins, formula=bill_depth_mm~bill_length_mm)

summary(billDepthAndLengthModel)
```

While this output is intimidating, we only need two numbers for our purposes: the estimate intercept (b) and the slope (m), located right below that number.

Our line of regression can be represented in the equation y = -0.08502x + 20.88574

Here is the same scatter plot with the line of best fit:

```{r, warning=FALSE}
ggplot(data=penguins, mapping=aes(x=bill_length_mm, y=bill_depth_mm))+
  geom_point()+
  geom_smooth(method="lm")
```
